Strong lensing selection effects

### 2022-08-09 paper/seleff.tex
unlensed detection

The most striking difference between the samples is in the stellar mass, which is much larger in the lenses than in the general population. Lenses tend to also have a larger halo mass, a smaller half-light radius at fixed stellar mass, and a larger stellar population synthesis mismatch parameter. The distribution in ellipticity and inner dark matter slope of the lenses instead look very similar to that of the parent population.
%This outcome was expected, given the results of the analysis of \Sref{sect:indlenses}.
Additionally, we can see that the amplitude of the strong lensing bias appears to be always larger in the lens sample with the more restrictive selection on Einstein radius.

>>>>

The most striking difference between the samples is in the stellar mass: strong lensing selects preferentially galaxies with larger values of $\mobs$.
%, which is much larger in the lenses than in the general population. 
Lenses tend to also have a larger halo mass, a smaller half-light radius at fixed stellar mass, and a larger stellar population synthesis mismatch parameter. The distribution in ellipticity and inner dark matter slope of the lenses instead look very similar to that of the parent population.
%This outcome was expected, given the results of the analysis of \Sref{sect:indlenses}.
Additionally, we can see that the amplitude of the strong lensing bias appears to be always larger in the lens sample with the more restrictive selection on Einstein radius.

\Fref{fig:sourcepars} shows the distribution in parameters describing the background source population, for the same simulations of \Fref{fig:lenspars}. 
The unlensed distribution (blue filled contours) was obtained by taking into account the detectability of the sources: we drew a large number of sources from the distribution described in section \ref{ssec:sources}, simulated images of them with the same properties as the images of the lenses, and selected only those galaxies detected with $S/N>10$.
Therefore, the resulting distribution is different to the parent distribution from which the simulation was drawn.

Since we want to quantify the effect of lensing with respect to the case with no lensing for a given survey, we took into account the detectability of the sources when determining the unlensed background galaxy distribution.

Strong lensing tends to preferentially select sources at higher redshift, especially in the more restrictive case with $\theta_{\mathrm{Ein,min}} > 1.0’’$.
This is because, at fixed lens properties, increasing the source redshift lowers the critical surface mass density and, consequently, increases the size of the caustics and the Einstein radius.
The distribution in source magnitude is also significantly affected by the strong lensing selection.
In a somewhat counter-intuitive way, the $\msource$ distribution of the lenses peaks at brighter magnitudes than the parent distribution. The parent sample, however, is simply the input of the simulations, and does not take into account whether sources are detectable or not. The detection limit that we assumed in our simulations is at $\msource\approx25$, and therefore the source magnitude distribution in the absence of lensing would drop rapidly. 
In the absence of lensing, this distribution would drop rapidly beyond the detection limit $\msource\approx25$. Lensing, however, allows the detection of sources at fainter magnitudes, in virtue of the magnification, and therefore the $\msource$ distribution shows a faint-end tail beyond $\msource=25$.
We can also see a bias in favour of sources with a very low S\'{e}rsic index. This is a consequence of the selection in magnitude: our source sample shows a strong correlation between $\msource$ and $n$: the brightest sources tend to have the smallest values of the S\'{e}rsic index.
